### 2014-04-08 paper.tex
initial commit

\documentclass[twocolumn]{article}
%\documentclass{article}
\usepackage{fullpage}
% you really only need to use one of graphicx or epsfig
% there are example of using each one in this file
%\usepackage[dvips]{graphicx}
%usepackage{epsfig}
\usepackage{pgfplots}

% if you want two columns per page, but you also need to reformat the figures
% to span both columns if you do this
%\twocolumn
\begin{document}
\title{Testing Characteristics of the Linux Page Replacement Policy}

\author{Callen Rain \& Justin Cosentino}

\maketitle

\begin{abstract}
A Brief summary of problem examining and results found
\end{abstract}

\section{Data}
%--------------------------------------------%

\begin{figure*}[t]
\begin{tikzpicture}
\begin{axis}[
    title = {MRU vs. LRU Test},
    width=8cm,
    xtick={1,...,2},
    xticklabels={1,2,3,4},
    grid=major,
    ybar, 
    bar width=15pt, 
    ylabel=Page Faults, 
    xlabel=Phase, 
    ymin=0, 
        enlarge x limits={abs=1}
    ]

\addplot[
    fill=blue!25,
    error bars/.cd,
        y dir=both,
        y explicit
] 
table [y error=error] {
x   y           error    label
1   8608.00 86.23 1
2   128.00 3.87 2
};

\addplot[
    fill=red!25,
    error bars/.cd,
        y dir=both,
        y explicit
] 
table [y error=error] {
x   y           error    label
1   7775.00   41.76 1
2   978.00   32.60 2
};
\legend{Case 1,Case 2}
\end{axis}
\end{tikzpicture}
\end{figure*}

\begin{table}[h]
\begin{center}
\begin{tabular}{|c|c|r|r|}
\hline
{\bf Phase } & {\bf Case} & {\bf Page Faults} & {\bf Error } \\
\hline
1 & 1 & 8608.00 & $\pm$ 86.23 \\
& 2 & 7775.00 & $\pm$ 41.76 \\
\hline
2 & 1 & 128.00 & $\pm$ 3.87 \\
& 2 & 978.00 & $\pm$ 32.60 \\
\hline
\end{tabular}
\caption{ \label{rawio} MRU versus LRU test data
{\em  Description  }}
\end{center}
\end{table}

%--------------------------------------------%
\begin{figure*}[h]
\begin{tikzpicture}
\begin{axis}[
    title = {FIFO vs. LRU Test},
    width=10cm,
    xtick={1,...,4},
    xticklabels={1,2,3,4},
    grid=major,
    ybar, 
    bar width=15pt, 
    ylabel=Page Faults, 
    xlabel=Phase, 
    ymin=0, 
    enlarge x limits={abs=1}
    ]

\addplot[
    fill=blue!25,
    error bars/.cd,
        y dir=both,
        y explicit
] 
table [y error=error] {
x   y           error    label
1   7708.00 158.27 1
2   976.00 20.27 2
3   976.00 8.89 3
4   0.00 0.00 4
};

\addplot[
    fill=red!25,
    error bars/.cd,
        y dir=both,
        y explicit
] 
table [y error=error] {
x   y           error    label
1   6832.00   128.12 1
2   993.00   20.12 2
3   979.00   33.21 3
4   983.00   15.59 4
};
\legend{Case 1,Case 2}
\end{axis}
\end{tikzpicture}
\end{figure*}

\begin{table}[h]
\begin{center}
\begin{tabular}{|c|c|r|r|}
\hline
{\bf Phase } & {\bf Case} & {\bf Page Faults} & {\bf Error } \\
\hline
1 & 1 & 7708.00 & $\pm$ 158.27 \\
& 2 & 6832.00 & $\pm$ 128.12 \\
\hline
1& 1 & 976.00 & $\pm$ 20.27 \\
& 2 & 993.00 & $\pm$ 20.12 \\
\hline
2 & 1 & 976.00 & $\pm$ 8.89 \\
& 2 & 979.00 & $\pm$ 33.21 \\
\hline
1 & 1 & 0.00 & $\pm$ 0.00 \\
& 2 & 983.00 & $\pm$ 15.59 \\
\hline
\end{tabular}
\caption{ \label{rawio} FIFO versus LRU test data
{\em  Description  }}
\end{center}
\end{table}

%--------------------------------------------%
\begin{figure*}[h]
\begin{tikzpicture}
\begin{axis}[
    title = {Working Set vs. LRU Test},
    width=10cm,
    xtick={1,...,3},
    xticklabels={1,2,3,4},
    grid=major,
    ybar, 
    bar width=15pt, 
    ylabel=Page Faults, 
    xlabel=Phase, 
    ymin=0, 
    enlarge x limits={abs=1}
    ]

\addplot[
    fill=blue!25,
    error bars/.cd,
        y dir=both,
        y explicit
] 
table [y error=error] {
x   y           error    label
1   3614.00 131.24 1
2   1.00 1.00 2
3   3990.00 24.58 3
};

\addplot[
    fill=red!25,
    error bars/.cd,
        y dir=both,
        y explicit
] 
table [y error=error] {
x   y           error    label
1   5142.00   320.20 1
2   1.00   0.00 2
3   2163.00   13.96 3
};
\legend{Case 1,Case 2}
\end{axis}
\end{tikzpicture}
\end{figure*}

\begin{table}[h]
\begin{center}
\begin{tabular}{|c|c|r|r|}
\hline
{\bf Phase } & {\bf Case} & {\bf Page Faults} & {\bf Error } \\
\hline
1 & 1 & 3614.00 & $\pm$ 131.24 \\
& 2 & 5142.00 & $\pm$ 320.20 \\
\hline
1& 1 & 1.00 & $\pm$ 1.00 \\
& 2 & 1.00 & $\pm$ 0.00 \\
\hline
2 & 1 & 3990.00 & $\pm$ 24.58 \\
& 2 & 2163.00 & $\pm$ 13.96 \\
\hline
\end{tabular}
\caption{ \label{rawio} Working Set versus LRU test data
{\em  Description  }}
\end{center}
\end{table}

%--------------------------------------------%

\section{Introduction}

Here are some font examples: {\em this is an example of how to get italics fonts } here is {\bf bold} here is a code font {\tt foo(int x);}.

If you want a new paragraph, just put a blank line between the 
current paragraph

and the next paragraph.

\noindent noindent will remove the auto indenting of paragraphs within
a section.

If you want a larger break between paragraphs end the line with slash slash
and add a blank line between it and the next paragraph\\

To refer to a figure tha appears anywhere in the document use 
{\tt backslash ref\{labelname\}} and note where the labelname is defined in
the figure (the tilda glues it to the preceding word of text).
For example: (see Figure~\ref{compvsinterp}).

Here is an example of how to incorporate a figure in a document...you need
to use one of the psfig or epsfig style files to include a postscript or 
an encapsulated postscript file.  These style files also include macros for
positioning your figure around the text, for changing the size of your figure
for adding captions and labels...

I can force a figure to the top or bottom of a page by using the [t] or [b].
I can also try to force it to come after the prose it follows in the .tex
file by using [!htb]:

The .cls and .sty files typically contain comments on how to use them in your latex
document...if you look at psfig.sty, the file contains comments that summarize
commands implemented by this style file and how to use them

Here is an example of how to print somthing using a courier font
function \texttt{invokeMethod} while interpreting AP method \texttt{foo}. The 
application program developer wants performance data that measures the

\section {Here is how to create an enumerated list:}

\begin{enumerate}
\item {\bf CPUtime, $<$/APCode/foo.class/foo$>$}:\\
  note the dollar signs around the above are interpreted as math mode.  In
  this mode some characters have special meaning that result in special
  formatting.  Latex is particularly good for representing mathematical
  functions via math mode.  You can also put a large block of text in math
  mode using begin and end.

\item  The next thing

\item {\bf methodCallTime, $<$/APCode/foo.class/foo$>$}:
  I used to have a mathematical definition for this.
\end{enumerate}

\subsection {Tables in Latex }

Tables are not the easiest interface to use, but you can have latex generate tables for you.
The first part {\tt e.g. |c|l|r|rr|} specifies the number of columns implicitly
and the formating for each column (centered, left, right).  If you put a 
vertical bar between entries, then a vertical bar will be drawn between the
columns otherwise not.  The first example below has 3 columns, the first
uses centered formatting that next two right.  Each row of values is
listed on a separate line.  Ampersands are used between each row's column 
values and backslash-backslah ends a row.   hline can be used to draw
horizontal lines.  multicolumn can be used to span columns and to
change the default column formatting.

Here is a larger example that also demostrates multicolumn directive:

\section {Some random prose to make this longer}

This course is a introduction to the theory, design, and implementation of operating systems. An operating system is the software layer between users and the computer hardware. It implements abstractions that are easier to program than the underlying hardware (e.g. processes, virtual memory, file systems), and it manages the machine's resources (e.g. memory, cpus, disk, network interfaces, and other devices). We will cover the following topics: processes (including synchronization, communication, and scheduling), memory (main memory allocation strategies, virtual memory, and page replacement policies), file systems (including naming and implementation issues), I/O (including devices, drivers, SSDs and disks, and disk scheduling), protection and security, distributed systems and other advanced topics. Prerequisites: CPSC 35 and CPSC 31 required.

This course is a broad introduction to computer science that focuses on how a computer works and how programs run on computers. We examine the hardware and software components required to go from a program expressed in a high-level programming language like C or Python to the computer actually running the program. This course takes a bottom-up approach to discovering how a computer works, and introduces parallel and distributed computing with a specific focus on parallelism for multicore and other shared memory systems. Topics include theoretical models of computation, data representation, machine organization, assembly and machine code, memory, I/O, the stack, the operating system, compilers and interpreters, processes and threads, and synchronization. In addition to parallel programming, we will discuss parallel computers and system-level support for parallel computing. Prerequisite: Completion of CS21 or its equivalent.

\subsection {Goals for the Course:}

\begin{itemize}
    \item To understand how a sequential or parallel program goes from being expressed in a high-level programming language to being run on the underlying system.
    \item To understand and analyze the systems costs associated with application performance.
    \item To understand the role of the operating system and some of the abstractions it implements to support efficiently running programs.
    \item To understand shared memory parallel computing and to be able to "think in parallel" algorithmically.
    \item To become proficient in using gdb and valgrind to debug and examine program execution state.
    \item To design and implement solutions to large problems using the C programming language.
    \item To design and implement parallel solutions to programming problems that require synchronization using pthreads. 
\end{itemize}

\section{More Random Prose}

This course covers a broad range of topics related to parallel and distributed computing, including parallel and distributed architectures and systems, parallel and distributed programming paradigms, parallel algorithms, and scientific and other applications of parallel and distributed computing. In lecture/discussion sections, students examine both classic results as well as recent research in the field. The lab portion of the course includes programming projects using different programming paradigms, and students will have the opportunity to examine one course topic in depth through an open-ended project of their own choosing. Course topics may include: multi-core, SMP, MMP, client-server, clusters, clouds, grids, peer-to-peer systems, GPU computing, scheduling, scalability, resource discovery and allocation, fault tolerance, security, parallel I/0, sockets, threads, message passing, MPI, RPC, distributed shared memory, data parallel languages, MapReduce, parallel debugging, and applications of parallel and distributed computing.

Class will be run as a combination of lecture and seminar-style discussion. During the discussion based classes, students will read research papers prior to the class meeting that we will discuss in class. During the first part of the course, we will examine different parallel and distributed programming paradigms. During the second part of the course, students will propose and carry out a semester-long research project related to parallel and/or distributed computing. The department's gigabit cluster, two eight processor workstations, as well as the CS lab machines, are available for course projects.

Previous course work in Operating Systems, Networking, Databases, or Architecture is helpful, but not necessary for taking this course.

\section{Conclusion}

To figure out how to do something in latex, look at some examples and some
on-line documentation, or google for it.  Often times it is very difficult
to get latex to format something differently from what its default formatting
is.

\end{document}

### 2014-04-09 paper.tex
edits

@@
 \begin{abstract}
-A Brief summary of problem examining and results found
+This paper describes some of the techniques used to identify the behavior of the Linux page replacement and discusses the implementation of a system call that provides page replacement statistics.
+The system call provides page fault statistics for minor and major faults for a single process, processes grouped by owners, and all processes on the system.
+Several tests were developed that accessed large segments of memory on a Linux virtual machine.
+In each of the tests, access patterns exploited some page replacement scheme and favored others.
+An initial hypothesis of a Least Recently Used (LRU) algorithm was partially supported, although there was some evidence that the Linux implementation has characteristics resembling a working set.
 \end{abstract}
 
-\section{Data}
+\section{Introduction}
+
+We began
+
+ Describe at a high-level what you are doing and why, and describe what your high-level hypothesis is (i.e.
+"We started out guessing that Linux's page replacement policy is X, but our experiments will show that it is in fact, more like Y" or "we start out making no assumptions about Linux's page replacement policy, and our experiments will answer the question of is it more like X, Y or Z.").
+
+\section{System Call}
+Short paragraph describing how you implemented your system call to get get per-process, per-owner, and system wide page fault counts?
+Also, describe other Unix utilities you used to collect performance data for your experiments.
+
+\section{Methodology}
+\subsection{Experiments}
+\subsubsection{Ruling Out Most Recently Used}
 %--------------------------------------------%
 
-\begin{figure*}[t]
+\begin{figure}[h]
 \begin{tikzpicture}
@@
     ybar, 
-    bar width=15pt, 
+    bar width=10pt, 
     ylabel=Page Faults, 
@@
 \end{tikzpicture}
-\end{figure*}
+\end{figure}
 
@@
 \caption{ \label{rawio} MRU versus LRU test data
-{\em  Description  }}
+{\em  Description DescriptionDescriptionDescriptionDescription  }}
 \end{center}
@@
 %--------------------------------------------%
-\begin{figure*}[h]
+
+\subsubsection{Ruling Out First In First Out}
+%--------------------------------------------%
+\begin{figure}[h]
 \begin{tikzpicture}
@@
     title = {FIFO vs. LRU Test},
-    width=10cm,
+    width=8cm,
     xtick={1,...,4},
@@
 %--------------------------------------------%
-\begin{figure*}[h]
+
+\subsubsection{Balance Between Working Set Least Recently Used}
+%--------------------------------------------%
+\begin{figure}[h]
 \begin{tikzpicture}
@@
     title = {Working Set vs. LRU Test},
-    width=10cm,
+    width=8cm,
     xtick={1,...,3},
@@
 
-\section{Introduction}
+\subsubsection{Random LRU Behavior}
 
-Here are some font examples: {\em this is an example of how to get italics fonts } here is {\bf bold} here is a code font {\tt foo(int x);}.
+Main part: describe your experiments and results: For each experiment, you should:
 
-If you want a new paragraph, just put a blank line between the 
-current paragraph
+Clearly state what your hypothesis is (i.e. "Application that does X should be a bad/good case for Linux's page replacement policy because ...)
 
-and the next paragraph.
+Explain how your experiment is testing that hypothesis (We are testing this hypothesis by running application P, which does ..., N times on and collecting X,Y, and Z metrics using our system call, vmstat, time,...
+With these measures, we can see whether or not ...
+because ...).
 
-\noindent noindent will remove the auto indenting of paragraphs within
-a section.
+Present your Results.
 
-If you want a larger break between paragraphs end the line with slash slash
-and add a blank line between it and the next paragraph\\
+Explain your results! 
+(Our results show that ...
+This does/does not match our expected result, because ...
+(if it doesn't, think about some other tests you could run to explain why it doesn't, or look for data you do have to explain it)
 
-To refer to a figure tha appears anywhere in the document use 
-{\tt backslash ref\{labelname\}} and note where the labelname is defined in
-the figure (the tilda glues it to the preceding word of text).
-For example: (see Figure~\ref{compvsinterp}).
-
-Here is an example of how to incorporate a figure in a document...you need
-to use one of the psfig or epsfig style files to include a postscript or 
-an encapsulated postscript file.  These style files also include macros for
-positioning your figure around the text, for changing the size of your figure
-for adding captions and labels...
-
-I can force a figure to the top or bottom of a page by using the [t] or [b].
-I can also try to force it to come after the prose it follows in the .tex
-file by using [!htb]:
-
-The .cls and .sty files typically contain comments on how to use them in your latex
-document...if you look at psfig.sty, the file contains comments that summarize
-commands implemented by this style file and how to use them
-
-Here is an example of how to print somthing using a courier font
-function \texttt{invokeMethod} while interpreting AP method \texttt{foo}. The 
-application program developer wants performance data that measures the
-
-\section {Here is how to create an enumerated list:}
-
-\begin{enumerate}
-\item {\bf CPUtime, $<$/APCode/foo.class/foo$>$}:\\
-  note the dollar signs around the above are interpreted as math mode.  In
-  this mode some characters have special meaning that result in special
-  formatting.  Latex is particularly good for representing mathematical
-  functions via math mode.  You can also put a large block of text in math
-  mode using begin and end.
-
-\item  The next thing
-
-\item {\bf methodCallTime, $<$/APCode/foo.class/foo$>$}:
-  I used to have a mathematical definition for this.
-\end{enumerate}
-
-\subsection {Tables in Latex }
-
-Tables are not the easiest interface to use, but you can have latex generate tables for you.
-The first part {\tt e.g. |c|l|r|rr|} specifies the number of columns implicitly
-and the formating for each column (centered, left, right).  If you put a 
-vertical bar between entries, then a vertical bar will be drawn between the
-columns otherwise not.  The first example below has 3 columns, the first
-uses centered formatting that next two right.  Each row of values is
-listed on a separate line.  Ampersands are used between each row's column 
-values and backslash-backslah ends a row.   hline can be used to draw
-horizontal lines.  multicolumn can be used to span columns and to
-change the default column formatting.
-
-Here is a larger example that also demostrates multicolumn directive:
-
-\section {Some random prose to make this longer}
-
-This course is a introduction to the theory, design, and implementation of operating systems.
-An operating system is the software layer between users and the computer hardware.
-It implements abstractions that are easier to program than the underlying hardware (e.g.
-processes, virtual memory, file systems), and it manages the machine's resources (e.g.
-memory, cpus, disk, network interfaces, and other devices).
-We will cover the following topics: processes (including synchronization, communication, and scheduling), memory (main memory allocation strategies, virtual memory, and page replacement policies), file systems (including naming and implementation issues), I/O (including devices, drivers, SSDs and disks, and disk scheduling), protection and security, distributed systems and other advanced topics.
-Prerequisites: CPSC 35 and CPSC 31 required.
-
-This course is a broad introduction to computer science that focuses on how a computer works and how programs run on computers.
-We examine the hardware and software components required to go from a program expressed in a high-level programming language like C or Python to the computer actually running the program.
-This course takes a bottom-up approach to discovering how a computer works, and introduces parallel and distributed computing with a specific focus on parallelism for multicore and other shared memory systems.
-Topics include theoretical models of computation, data representation, machine organization, assembly and machine code, memory, I/O, the stack, the operating system, compilers and interpreters, processes and threads, and synchronization.
-In addition to parallel programming, we will discuss parallel computers and system-level support for parallel computing.
-Prerequisite: Completion of CS21 or its equivalent.
-
-\subsection {Goals for the Course:}
-
-\begin{itemize}
-    \item To understand how a sequential or parallel program goes from being expressed in a high-level programming language to being run on the underlying system.
-    \item To understand and analyze the systems costs associated with application performance.
-    \item To understand the role of the operating system and some of the abstractions it implements to support efficiently running programs.
-    \item To understand shared memory parallel computing and to be able to "think in parallel" algorithmically.
-    \item To become proficient in using gdb and valgrind to debug and examine program execution state.
-    \item To design and implement solutions to large problems using the C programming language.
-    \item To design and implement parallel solutions to programming problems that require synchronization using pthreads. 
-\end{itemize}
-
-\section{More Random Prose}
-
-This course covers a broad range of topics related to parallel and distributed computing, including parallel and distributed architectures and systems, parallel and distributed programming paradigms, parallel algorithms, and scientific and other applications of parallel and distributed computing.
-In lecture/discussion sections, students examine both classic results as well as recent research in the field.
-The lab portion of the course includes programming projects using different programming paradigms, and students will have the opportunity to examine one course topic in depth through an open-ended project of their own choosing.
-Course topics may include: multi-core, SMP, MMP, client-server, clusters, clouds, grids, peer-to-peer systems, GPU computing, scheduling, scalability, resource discovery and allocation, fault tolerance, security, parallel I/0, sockets, threads, message passing, MPI, RPC, distributed shared memory, data parallel languages, MapReduce, parallel debugging, and applications of parallel and distributed computing.
-
-Class will be run as a combination of lecture and seminar-style discussion.
-During the discussion based classes, students will read research papers prior to the class meeting that we will discuss in class.
-During the first part of the course, we will examine different parallel and distributed programming paradigms.
-During the second part of the course, students will propose and carry out a semester-long research project related to parallel and/or distributed computing.
-The department's gigabit cluster, two eight processor workstations, as well as the CS lab machines, are available for course projects.
-
-Previous course work in Operating Systems, Networking, Databases, or Architecture is helpful, but not necessary for taking this course.
+Quality of Experiments is much, much better than quantity; a few well thought out and well done experiments is sufficient.
 
@@
 
-To figure out how to do something in latex, look at some examples and some
-on-line documentation, or google for it.  Often times it is very difficult
-to get latex to format something differently from what its default formatting
-is.
+It should include a statement of the major result(s) of your experiments (which policy(ies) is Linux's most like?
+and in what way(s)?).
+Also, tell me what you found to be the most difficult part of the hypothesis testing phase of this assignment.
 

### 2014-04-09 paper.tex
edits

@@
 
-We began
-
- Describe at a high-level what you are doing and why, and describe what your high-level hypothesis is (i.e.
-"We started out guessing that Linux's page replacement policy is X, but our experiments will show that it is in fact, more like Y" or "we start out making no assumptions about Linux's page replacement policy, and our experiments will answer the question of is it more like X, Y or Z.").
+We began with a hypothesis that the Linux system uses a Least Recently Used (LRU) page replacement algorithm, based on our analysis of the relevant kernel code and our numerous attempts in class to approximate this algorithm.
+Our first few tests confirmed that the replacement policy is more like LRU than like Most Recently Used (MRU) or First In First Out (FIFO), but our test to find some distinction between LRU and a Working Set (WS) algorithm did lean in favor of some Working Set behavior rather than strict LRU.
 
 \section{System Call}
-Short paragraph describing how you implemented your system call to get get per-process, per-owner, and system wide page fault counts?
-Also, describe other Unix utilities you used to collect performance data for your experiments.
+Our experimental methodology relies heavily on the page fault statistics system call that we implemented for the first part of this lab.
+The system call reports the number of page faults for all processes in the system, all the processes being run by the owner of the current process, or just the page faults for the current process.
+For each of these operations, the system call reports the major faults, minor faults, and the number of processes being reported on.
+We collected statistics primarily on major faults, as they are the page faults that result from having to replace pages in memory.
+
+The system call first does all the normal error checking, ensuring that the pointer passed in is writable and valid, that the process or processes requested exist, and that the flag is valid.
+We defined the ${\tt pf\_info\_struct()}$
+and the included flags in a header file.
+If the user enters a pid of -1, we used the ${\tt for\_each\_process()}$ macro to loop over all the processes, incrementing the major faults, minor faults, and number of processes for each one.
+We then fill the fields of the included struct.
+If the pid is positive, we check the inputted flag.
+If the flag is ${\tt PGFLTSTAT\_PROC}$, we do the same procedure for just the current process.
+If the flag is ${\tt PGFLTSTAT\_OWNER}$, we look for the owner id of the current process, and loop through all the processes again, collecting data only for the ones that share the uid.
 

### 2014-04-09 paper.tex
edits
added methology subsection

@@
 \subsection{Experiments}
+
+We carried out four experiments, all contained as subroutines within one file ({\tt mem.c}).
+The program selects one of the subroutines from a command line argument.
+In every test, the Virtual Machine was configured to have $256$MB of physical memory and $128$ buckets of unsigned ints (4B) were allocated.
+Thus each bucket had 4MB of data contained with it.
+Depending on the availability of memory after the boot, the system could hold slightly less than 64 of these buckets.
+We used this fact to plan some of our tests.
+
+Each test (except the random LRU test) involves two cases.
+In either case for a specific test, the system does roughly the same allocation routine with a slightly different final phase.
+In these last phases, we attempted to exploit some characteristic of the page replacement algorithm.
+In one case, we would access memory that we predicted would be replaced by the algorithm, and in the other case, we would access memory that we guessed had not been replaced.
+By printing out major fault statistics between the phases, we could get an idea of whether our hypothesis was correct or not by calculating the average page faults for each phase.
+We report this average over the number of trials and include the standard deviation as a measure of the error.
+
 \subsubsection{Ruling Out Most Recently Used}

### 2014-04-09 paper.tex
changes

@@
 Several tests were developed that accessed large segments of memory on a Linux virtual machine.
-In each of the tests, access patterns exploited some page replacement scheme and favored others.
-An initial hypothesis of a Least Recently Used (LRU) algorithm was partially supported, although there was some evidence that the Linux implementation has characteristics resembling a working set.
+In each of the tests, access patterns exploited specific page replacement schemes and favored others.
+An initial hypothesis of a Least Recently Used (LRU) algorithm was partially supported, although there was some evidence that the Linux implementation is a combination of various algorithms and has some characteristics resembling a working set.
 \end{abstract}
@@
 We began with a hypothesis that the Linux system uses a Least Recently Used (LRU) page replacement algorithm, based on our analysis of the relevant kernel code and our numerous attempts in class to approximate this algorithm.
-Our first few tests confirmed that the replacement policy is more like LRU than like Most Recently Used (MRU) or First In First Out (FIFO), but our test to find some distinction between LRU and a Working Set (WS) algorithm did lean in favor of some Working Set behavior rather than strict LRU.
+Our first few tests confirmed that the replacement policy is more like LRU than like Most Recently Used (MRU) and First In First Out (FIFO), but our test to find some distinction between LRU and a Working Set (WS) algorithm did lean in favor of some Working Set behavior rather than strict LRU.
+It is evident that the Linux page replacement implementation is not strictly one of the algorithms outline in class, but rather exhibits characteristics of both WS and LRU.
 
@@
 
-The system call first does all the normal error checking, ensuring that the pointer passed in is writable and valid, that the process or processes requested exist, and that the flag is valid.
-We defined the ${\tt pf\_info\_struct()}$
-and the included flags in a header file.
+The system call first does normal error checking, ensuring that the pointer passed in is writable and valid, that the process or processes requested exist, and that the flag is valid.
+We defined the ${\tt pf\_info\_struct}$, which stores the  major faults, minor faults, and the number of associated processes, and the included flags in a header file.
 If the user enters a pid of -1, we used the ${\tt for\_each\_process()}$ macro to loop over all the processes, incrementing the major faults, minor faults, and number of processes for each one.
@@
 If the flag is ${\tt PGFLTSTAT\_OWNER}$, we look for the owner id of the current process, and loop through all the processes again, collecting data only for the ones that share the uid.
+Once completed, this information is written to the provided ${\tt pf\_info\_struct}$.
 
@@
 Depending on the availability of memory after the boot, the system could hold slightly less than 64 of these buckets.
-We used this fact to plan some of our tests.
+We used this fact to plan some of our tests and memory access patterns.
 
@@
 \subsubsection{Ruling Out Most Recently Used}
+As discussed in class, the Most Recently Used (MRU) strategy is often a poor choice for a page replacement algorithm.
+Since there are only a few application specific cases in which this algorithm works well and Linux handles such a diverse workload, we hypothesize that MRU is not used as Linux's page replacement policy.
+
+We test this hypothesis by running each case of the first experiment in file {\tt mem.c}.
+In both cases, this subroutine first writes 128 4MB buckets to a block of memory, which is significantly larger than what can fit in RAM.
+We now access a small chunk (8-12 buckets) of this large block in one of two ways.
+In the first case, we read in the last 12 buckets.
+Were the Linux page replacement policy a MRU algorithm, we would expect these memory accesses to cause major page faults as recently-used page entries would be evicted once the page table filled up.
+On the other hand, were the page replacement policy strictly LRU, we would expect these entries to still be within the page table and minimal major page faulting.
+As seen in Figure 1 and Table 1, this experiment resulted in very few major page faults, supporting our hypothesis that MRU is not used as Linux's page replacement policy.
+
+In the second case, we read in the first 8 buckets of the large block of memory.
+Were the Linux page replacement policy a MRU algorithm, we would expect these memory accesses to cause very few major page faults as these least recently-used page entries would be less likely to be evicted than their recent counterparts.
+Conversely, were Linux page replacement policy a LRU algorithm we would expect major page faults to occur as these least recently-used entries would have been evicted once the page table filled up.
+The number of major page faults that occurred in this case, as depicted in Figure 1 and Table 1, were nearly an order of magnitude greater than those that occurred in the first case, further supporting our initial hypotheses.
+Not only does this experiment refute the possibility that Linux implements a MRU page replacement algorithm, but the displayed behavior implies an LRU implementation.
+
 %--------------------------------------------%
 
-\begin{figure}[h]
+\begin{figure}[h!]
 \begin{tikzpicture}
@@
 
-\begin{table}[h]
+\begin{table}[h!]
 \begin{center}
@@
 \subsubsection{Ruling Out First In First Out}
+Given that First In First Out (FIFO) page replacement policies do not take into account memory usage information and results in Bélády's anomaly, we hypothesize that Linux does not make us of a FIFO based algorithm.
+We confirm this hypothesis by running the second subroutine in our experiment file.
+In both cases of this experiment, the application first writes 128 4MB buckets to a block of memory and then repeatably reads and writes to the first 8 buckets.
+Next, the application writes to the last 8 buckets once, effectively removing the first 8 buckets if Linux utilizes a FIFO algorithm.
+In the first case, this subroutine then reads the first 8 buckets.
+Were the Linux page replacement policy a FIFO algorithm, we would expect these read operations to major page fault, as they were the first entries related to this memory block to be added to the page table and would have been removed after the last write.
+On the other hand, if it is an LRU implementation we would expect minimal page faults to occur as this memory was recently accessed.
+As seen in Figure 2 and Table 2, no major page faults occurred, supporting our hypothesis and further strengthening our claim that Linux implements an LRU page replacement algorithm.
+
+In the second case, the last phase of this application reads buckets 8-16.
+We would expect a FIFO algorithm not to page fault when performing these operations, as these buckets would not yet have been evicted from the page table.
+In reality, these operations resulted in nearly 1000 major page faults.
+This case also implies that the Linux page replacement algorithm is not a FIFO based algorithm.
+
 %--------------------------------------------%
@@
 \hline
-1& 1 & 976.00 & $\pm$ 20.27 \\
+2& 1 & 976.00 & $\pm$ 20.27 \\
 & 2 & 993.00 & $\pm$ 20.12 \\
 \hline
-2 & 1 & 976.00 & $\pm$ 8.89 \\
+3 & 1 & 976.00 & $\pm$ 8.89 \\
 & 2 & 979.00 & $\pm$ 33.21 \\
 \hline
-1 & 1 & 0.00 & $\pm$ 0.00 \\
+4 & 1 & 0.00 & $\pm$ 0.00 \\
 & 2 & 983.00 & $\pm$ 15.59 \\
@@
 \end{table}
-
 %--------------------------------------------%
@@
 \subsubsection{Balance Between Working Set Least Recently Used}
+The first two experiments imply that Linux implements neither a MRU nor a FIFO page replacement algorithm.
+
 %--------------------------------------------%
@@
 \hline
-1& 1 & 1.00 & $\pm$ 1.00 \\
+2& 1 & 1.00 & $\pm$ 1.00 \\
 & 2 & 1.00 & $\pm$ 0.00 \\
 \hline
-2 & 1 & 3990.00 & $\pm$ 24.58 \\
+3 & 1 & 3990.00 & $\pm$ 24.58 \\
 & 2 & 2163.00 & $\pm$ 13.96 \\

### 2014-04-10 paper.tex
add graph

@@
 
+%--------------------------------------------%
+\begin{figure}[h]
+\begin{tikzpicture}
+\begin{axis}[
+    title = {Random vs. Sequential Access},
+    width=8cm,
+    xtick={1, 2},
+    xticklabels={1, 2},
+    grid=major,
+    ybar, 
+    bar width=10pt, 
+    ylabel=Page Faults, 
+    xlabel=Case, 
+    ymin=0, 
+    enlarge x limits={abs=1}
+    ]
+
+\addplot[
+    fill=red!25,
+    error bars/.cd,
+        y dir=both,
+        y explicit
+] 
+table [y error=error] {
+x   y           error    label
+1   9862.00   220.40 1
+2   4789.00   713.83 2
+};
+\end{axis}
+\end{tikzpicture}
+\end{figure}
+
+\begin{table}[h]
+\begin{center}
+\begin{tabular}{|c|c|r|r|}
+\hline
+{\bf Phase } & {\bf Case} & {\bf Page Faults} & {\bf Error } \\
+\hline
+1 & 1 & 9862.00 & $\pm$ 220.40 \\
+& 2 & 4789.00 & $\pm$ 713.83 \\
+
+\hline
+\end{tabular}
+\caption{ \label{rawio} Working Set versus LRU test data
+{\em  Description  }}
+\end{center}
+\end{table}
+
+%--------------------------------------------%
+
 Main part: describe your experiments and results: For each experiment, you should:
